### 2014-06-04 writing/S'14 paper/recursion.tex
Revision to quantifiers section
Intermediate paper update.

\section{Recursive structure}

@@

Recursive structure is an prominent feature of natural languages and
an important part of their expressivity, as it allows people to use
and interpret complex structures they have never encountered before.
Theoretical accounts of natural language syntax and semantics
therefore rely heavily on recursive structures, and we expect that
accurate computational models will have to come to grips with it as
well. The present section pursues this question for our two classes of
RNN. In evaluating the model, we take advantage of the fact that
recursive structures of this kind define potentially infinite
languages, by testing the model on strings that are longer and more
complex than any seen in testing.

@@

As in Section~\ref{sec:join}, we test this phenomenon within the
framework of natural logic, but we now replace the unanalyzed symbols
from that experiment with expressions that involve recursive
structure. The result is a truth-functionally complete classical
propositional logic: each atomic variable denotes either $\True$ or
$\False$, and the only operators are truth-functional ones.
Table~\ref{tab:pl} defines this logic, and Table~\ref{tab:plexs} gives
some examples of natural logic relational statements that we can make
in these terms. To compute these relations between statements, we
exhaustively enumerate the sets of assignments of truth values to
proposition variables that would satisfy each of the statements and
then convert the set-theoretic relation between those assignments into
one of the seven relations in Table~\ref{b-table}. This is equivalent
to proving theorems about these statements model-theoretically.

\newcommand{\plneg}{\mathop{\textit{not}}}
\newcommand{\pland}{\mathbin{\textit{and}}}
\newcommand{\plor}{\mathbin{\textit{or}}}

@@

% NOTE: There was some confusion at CSLI and the 224U lecture about
% whether this was theorem proving or some kind of model checking.
% Hopefully this is explicit enough.
Socher et al.~\cite{socher2012semantic} demonstrate the learning of a
logic in a matrix-vector RNN model somewhat similar to our own, but
the logic discussed here is substantially stronger, and a much better
approximation of the kind of structure that is needed for natural
language. The logic learned in that experiment is boolean, wherein the
atomic symbols are simply the values $0$ and $1$, rather than
variables over those values. While learning the operators of that
logic is not trivial, the outputs of each operator can be represented
accurately by a single bit. The statements of propositional logic
learned here describe conditions on the truth values of propositions
where those truth values are not known. As opposed to the two-way
contrasts seen in \cite{socher2012semantic}, this logic distinguishes
between $2^{6} = 64$ possible assignments of truth values, and
expressions of this logic define arbitrary conditions on these
possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
possible statements that the recursive model needs to be able to
distinguish.

For our experiments, we randomly generate large sets of syntactically
unique formulae and compute the natural logic relation between them.
We discard pairs in which either statement is a tautology or
contradiction (a statement that is true of either all or no possible
assignments), for which none of the seven relation labels in
Table~\ref{b-table} can hold. The resulting set of formula pairs is
then partitioned into bins based on the number of logical operators in
the longer of the two formulae. We then randomly sample 15\% of each
bin for a held out test set.

@@

% We deduplicate the data and discard pairs in which either statement is
% a tautology or contradiction (a statement that is true of either all
% or no possible assignments), for which none of the seven relation
% labels can accurately apply. We then divide the generated pairs into
% size bins based on the number of logical operators (\ii{and}, \ii{or},
% or \ii{not}) in the larger of the two pairs being compared, and
% discard examples of size greater than twelve by this measure. Finally,
% we randomly sample 15\% of each bin for a held out test set.

@@

\begin{figure}[htp]
  \centering
  \includegraphics[width=5.5in]{recursion\string_results.eps}
  \caption{Model performance by expression size.}  
  \label{prop-results}
\end{figure}

>>>>

\section{Recursive structure}\label{sec:recursion}

@@

Recursive structure is an prominent feature of natural languages and
an important part of their expressivity, as it allows people to use
and interpret complex structures they have never encountered before.
Theoretical accounts of natural language syntax and semantics
therefore rely heavily on recursive structures, and we expect that
accurate computational models will have to come to grips with them as
well. The present section pursues this question for our two classes of
RNN. In evaluating the model, we take advantage of the fact that
recursive structures of this kind define potentially infinite
languages, by testing the model on strings that are longer and more
complex than any seen in testing.

@@

As in Section~\ref{sec:join}, we test this phenomenon within the
framework of natural logic, but we now replace the unanalyzed symbols
from that experiment with expressions that involve recursive
structure. These expressions represent a truth-functionally complete classical
propositional logic: each atomic variable denotes either $\True$ or
$\False$, and the only operators are truth-functional ones.
Table~\ref{tab:pl} defines this logic, and Table~\ref{tab:plexs} gives
some examples of relational statements that we can make
in these terms. To compute these relations between statements, we
exhaustively enumerate the sets of assignments of truth values to
proposition variables that would satisfy each of the statements and
then convert the set-theoretic relation between those assignments into
one of the seven relations in Table~\ref{b-table}. As a result, each relational
statement represents a valid theorem of the propositional logic.

@@

% NOTE: There was some confusion at CSLI and the 224U lecture about
% whether this was theorem proving or some kind of model checking.
% Hopefully this is explicit enough.
Socher et al.~\cite{socher2012semantic} demonstrate the learning of a
logic in a matrix-vector RNN model somewhat similar to our own, but
the logic discussed here is substantially stronger, and a much better
approximation of the kind of structure that is needed for natural
language. The logic learned in that experiment is boolean, wherein the
atomic symbols are simply the values $0$ and $1$, rather than
variables over those values. While learning the operators of that
logic is not trivial, the outputs of each operator can be represented
accurately by a single bit. The statements of propositional logic
learned here describe conditions on the truth values of propositions
where those truth values are not known. As opposed to the two-way
contrasts seen in \cite{socher2012semantic}, this logic distinguishes
between $2^{6} = 64$ possible assignments of truth values, and
expressions of this logic define arbitrary conditions on these
possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
possible statements that the intermediate vector representations need 
to be able to distinguish.

For our experiments, we randomly generate a large set of  unique pairs 
of formulae and compute the relation that holds for each pair.
We discard pairs in which either statement is a tautology or
contradiction (a statement that is true of either all or no possible
assignments), for which none of the seven relation labels in
Table~\ref{b-table} can hold. The resulting set of formula pairs is
then partitioned into bins based on the number of logical operators in
the longer of the two formulae. We then randomly sample 15\% of each
bin for a held out test set.

@@

@@

\begin{figure}[htp]
  \centering
  \includegraphics[width=5.5in]{recursion\string_results.eps}
  \caption{Model performance on propositional logic, by expression size.}  
  \label{prop-results}
\end{figure}
